\documentclass[12pt]{article}
\usepackage{amsthm,amsmath,amsfonts,amsthm,amstext,amssymb,fullpage,framed,fancybox,graphicx,color,mdwlist,pifont, hyperref}
%\usepackage{fullpage}
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;} 
\usepackage[margin=1cm]{geometry}
\pagestyle{empty}
\newtheorem*{theorem*}{Theorem}

\newcommand{\set}[1]{\left\{ #1 \right\}}
\renewcommand{\neg}{\sim}
\newcommand{\st}{\text{ s.t. }}
\begin{document}
\centerline{\bf\large Worksheet for Week 4}

\vspace{25pt}


Note: Order of the quantifiers matters.

\begin{enumerate}

\item Consider the following statements and discuss which ones are true.

\begin{itemize}
\item $\forall m\in\mathbb Z, \exists n\in\mathbb Z$ s.t. $m^2<n$.
\item $\exists m\in\mathbb Z$ s.t. $\forall n\in\mathbb Z$, $m<n$.
\item $\exists m,n\in\mathbb Z$, s.t. $m^2<n$.
\item $\forall m,n\in\mathbb Z$, $m^2<n$.
\end{itemize}

\item Negate the statements below and determine whether the statements or their negations are true.

\begin{itemize}
\item $\forall x\in\mathbb R$, $\exists y\in\mathbb R$, s.t. $y+5=x$.
\item $\forall x\in\mathbb R$, $\exists n\in\mathbb N$, s.t. $x^n>0$.
\item $\forall n\in\mathbb R$, $(n^2\text{ is even})\implies (n \text{ is even})$.
\item $\exists y\in\mathbb R$, s.t. $\forall x\in\mathbb R$, $x^2\geq y$.
\item $\exists x\in\mathbb R$, s.t. $\forall y\in\mathbb R$, $(y\leq x-1)\implies (y^2-x^2\geq 4)$.
\end{itemize}

\item Prove the following statement:
\[ \forall x\in\mathbb R, \forall y\in\mathbb R, (\forall z>0, |x-y|<z)\implies (x=y).  \]
Hint: How can you rewrite the statement?

\item Let $\mathbb I$ denote the set of irrationals. Prove that $\forall x\in\mathbb I$, $(k\in\mathbb Q\implies \dfrac{x}{k}\in\mathbb I)$.

Now, use this result, and the fact that $\pi\in\mathbb I$ to show
\[\forall n\in\mathbb N, \exists y\in\mathbb I \st x>y>0.\]

\item Let $A$ be a set. We say that a function $f: A\to \mathbb R$ is bounded if $\exists m\in\mathbb R, \st \forall x\in A,\  |f(x)|\leq M$.

Use this definition to show that $f:\mathbb R\to \mathbb R$ defined as $f(x)=3x-7$ is not bounded.

Hint: If the statement above is the definition of a bounded function, what should a function that is NOT bounded satisfy?

\item Fact: $\forall x\in\mathbb R$, $\exists n\in\mathbb N$, s.t. $n\geq x$.

Use this fact to show
\[\forall \epsilon>0, \exists N\in\mathbb N \st \forall n\in\mathbb N,\ \dfrac{1}{n}<\epsilon.\] 
Hint: For $\epsilon, n>0$, we have $(\frac 1n<\epsilon) \iff (\frac{1}{\epsilon}<n)$.

\end{enumerate}

Before next week, watch videos 16 and 17 in \url{https://personal.math.ubc.ca/~PLP/auxiliary.html}.

\end{document} 